---
title: "Introduction to correctR"
author: "Trent Henderson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to correctR}
%\VignetteEngine{knitr::knitr}
%\usepackage[UTF-8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.height = 7,
  fig.width = 9
)
```

```{r, message = FALSE, warning = FALSE}
library(correctR)
```

`correctR` is a lightweight package that implements a small number of corrected test statistics for cases when samples of two machine learning model metrics (e.g., classification accuracy) are not independent (and therefore are correlated), such as in the case of resampling and $k$-fold cross-validation. We demonstrate the basic functionality here using some trivial examples.

## Setup

In the real world, we would have properly generated results from two models through random subsampling, $k$-fold cross-validation, or repeated $k$-fold cross-validation. However, for simplicity here, we are just going to simulate three datasets so we can get to the package functionality cleaner and easier. We are going to assume we are in a classification context and generate classification accuracy values.

```{r, message = FALSE, warning = FALSE}
# Data for random subsampling correction

x

# Data for k-fold cross-validation correction

x

# Data for repeated k-fold cross-validation correction

x
```

## Package functionality

x
